Similar Sentence Detection Using Locality Sensitive Hashing Technique on Wikipedia

### 2013-10-27 article/article-draft.tex
New draft of article

\title{Similar Sentence Detection Using Locality Sensitive Hashing Technique on Wikipedia\titlenote{This effort is part of the CMSC828G, Data Intensive Computing with MapReduce, class project.}}

@@

\author{
\alignauthor Samet Ayhan\\
       \affaddr{University of Maryland}\\
       \affaddr{Dept. of Computer Science}
       \email{sayhan@cs.umd.edu}
\alignauthor Joshua Bradley\\
       \affaddr{University of Maryland}\\
       \affaddr{Dept. of Computer Science}
       \email{jgbrad1@cs.umd.edu}
\alignauthor Sarah Weissman\\
       \affaddr{University of Maryland}\\
       \affaddr{College of Information Studies}\\
       \email{sew@umd.edu}
}

\date{May 9, 2013}

\maketitle
\begin{abstract}
The growth of Internet has enabled collaboration and cooperation on a large scale, resulting in an abundant number of near-duplicate web documents. The range of near-duplicate occurrences is even more evident within Wikipedia articles, due to its editing largely being open to the online community. Except for particularly sensitive and/or vandalism-prone pages that are "protected" to some degree from editing, the reader of an article can edit the text by copying from another article without prior approval. There is no current automated mechanism for truth validation. Although we are not inspired to measure the quality of Wikipedia articles, we are interested in finding near-duplicate occurrences at various granularity levels.

In this paper, we describe a novel similarity detection algorithm that utilizes a locality sensitive hashing (LSH) technique on the MapReduce framework. The algorithm has been designed and implemented to detect similar articles using large Wikipedia dumps. Experimental results appear to support that our method is able to efficiently and effectively detect similar articles across Wikipedia.
\end{abstract}

@@

\section{Introduction}
The online encyclopedia Wikipedia is a multilingual, free resource with its 26 million articles. All these articles, with over 4.2 million in English alone, are written collaboratively by volunteer contributors around the world. Most articles can be edited by anyone with internet access, thus there are currently approximately 100,000 active contributors. In light of these facts, it is clear that the unprecedented example of such a large-scale collaboration effort may result in duplicate or near duplicate articles at various granular levels, such as sentence-level or multi-gram granularity \cite{wiki:weblink}

Furthermore, Wikipedia has a history of generating controversy over editorial quality and factual correctness. Although some studies have found Wikipedia's accuracy to rival that of traditional encyclopedias, other studies have found numerous factual errors. Since Wikipedia may be edited anonymously, information is often freely copied between web sources and even between Wikipedia articles. Additionally, many articles in Wikipedia lack citations. There is little mechanism for detecting and correcting such errors. One potential source of error on Wikipedia is factual information that varies over time and is not updated. Wilkinson found that the distribution of article edits has a long tail, meaning that a small number of articles have many edits while most articles have few, and that the number of edits is directly related to article quality. Articles with low edits and low editorial attention, are less likely to be updated when new information is available \cite{wilkinson:wiki}.

As a step toward identifying factual errors, we investigated the problem of finding near duplicate sentences on Wikipedia. Such near duplicate sentences may reflect factual information that has been updated in one article and not another, or merely inconsistencies that should be flagged and addressed.

In order to come up with a novel algorithm that would allow us to detect near-duplicate sentences within Wikipedia articles, we decided that we needed to answer the following questions, first:
\begin{itemize}[noitemsep,nolistsep]
\item How effectively does Minhashing approximate edit distance? Edit distance seems like a reasonable metric for near duplicate sentences, however LSH does not map naturally to the edit distance metric.
\item What is the appropriate granularity for analysis? Is comparing multiple consecutive sentences more effective than comparing single sentences? Similarly, what is the best shingle size?
\item How well do techniques for detecting near duplicate documents work when applied at the sentence level?
\item Do near duplicate sentences correspond to factual errors? Can these errors be automatically detected and/or corrections suggested?
\end{itemize}

Along with these questions, we also considered the scalability and efficiency issues due to the large size of Wikipedia. To address these issues, we used the MapReduce computing framework and LSH techniques to group sentences with the same signatures and identify near duplicates.

@@

\section{Related Work}
A number of techniques have been implemented to identify similarity between documents in the literature. These techniques usually differ in terms of corpus of interest, signature representing the document, feature-set determined by the system, and the end goal. In addition to these differences, one can also consider the efficiency and scalability of the system, overall.

In this section, we will go over some of theses techniques offered in the literature. Additionally, we will highlight our novel contribution.

@@

In order to handle large Wikipedia dumps with the order of tens or hundreds of gigabytes and even a few terabytes, we used MapReduce framework in this work, that was introduced by Dean and Ghemawat \cite{dean:mapreduce}. The implementation is able to stream the compressed bz2 dumps and feed into mappers and reducers for further processing, which will be detailed in the following sections.

@@

%In order to measure document similarity we use a common LSH technique known as minhashing \cite{}. A minhash signature on a text document is calculated using a parametrized family of hash functions $F_i$, $1 \le i \le N$. In our case input documents are sentences. Each sentence is broken up into n-gram ``shingles''. For each shingle set $S$ we calculate $\{min_{s \in S}(F_i(s)\}$, the set of minimum hashes over the hash family. The signature on a document d is a vector of $K$ minhashes from the set of minimum hashes. In order to minimize false negatives we apply the technique sometimes known as ``banding'' \cite{Ullman} where multiple signatures are produced for each input document.

In order to measure document similarity we use a common LSH technique known as MinHash (\cite{broder:resemblance}). A minhash signature on a text document is calculated using a parametrized family of hash functions $F_i$, $1 \le i \le N$. In our case, we denote ''documents'' to mean sentences found within the content of Wikipedia articles. Each sentence is broken up into n-gram ``shingles''. For each shingle set $S$ we calculate the set $\{min_{s \in S}(F_i(s)\}$ of minimum hashes over the hash family. The signature on a document $d$ is represented as a vector of $K$ minhashes, chosen from the set $\{min_{s \in S}(F_i(s)\}$ of minimum hashes. In order to minimize false negatives we apply a technique sometimes known as ``banding'' \cite{ullman:massive} where multiple signatures are produced for each input document.

@@

Our implementation is built on top of the Apache Hadoop MapReduce framework, using utilities from the Cloud9 library (\url{https://github.com/lintool/Cloud9}) for parsing Wikipedia page text from the Wikipedia XML dump format. Sentences are parsed out over each document using a regular expression. Sentences that are too large or too small are discarded. The family of hash functions is implemented using a ``Multiply Shift'' (\url{http://en.wikipedia.org/wiki/Universal_hashing}) hashing scheme and is generated from a random seed using the Java Random class. The hash family is also parametrized by hash output key size and this value can be configured to affect the size of the output signatures.

@@

Jaccard similarity is a measure of set similarity while our stated goal is to find sentences in Wikipedia that have been copied from one article to another and possibly edited. A more natural measure to use for this application is edit distance, but there is no known good locality sensitive hashing family for edit distance \cite{Kirsch06distance-sensitivebloom}.

@@

In order to estimate the affect of shingle length on false positive rate we ran a number of experiments for a 1 G collection of wikipedia documents. For fixed $N=10$ we ran experiments for three values of $K=\{8,9,10\}$ and $5 \le L \le 15$. Setting a normalized edit distance threshold of $e = .25$, we calculated the total number of ''good'' ($e <= .25$) and ''bad'' ($e > .25$) (unique) sentence pairs as well as the false positive rate (\# bad pairs divided by the total number of sentences). Results are presented in Figure \ref{ed-grid}.

We can see from the graphs in Figure \ref{ed-grid} that all three statistics hit a low at shingle length 12 or 13. This means that in order to achieve the lowest false positive rate we should use parameter settings around $N=10, K=10$ and $L=12$ or $L=13$. However the number of \emph{good} pairs also hits a minimum at $L = 12$, which means that while overall the output is more precise, we are also missing output pairs.

Another feature of the graphs in Figure \ref{ed-grid} that stands out is that while the false positive rate and total number of bad pairs bottoms out at around $L=12$, for $L > 12$ the output quality actually decreases. This is counter to our intuition that output quality increases as the length of the shingle increases. One would expect larger shingles to correspond to closer edit distance scores, since larger shingles correspond to longer consecutive sequences of identical text. One explanation for these counter-intuitive results could be limitations of the underlying hash family. If longer shingle lengths lead to greater hash collisions this could explain the increased numbers of bad pairs. Improving the hash function to reduce collisions could improve results.

Since minimizing the false positive rate also means minimizing reducing good pairs, another option is to let parameters to maximize good output and use edit distance as a secondary filter. With this approach our experimental results suggest that the best parameter combination would be $N=10, K=8, L=10$.

\begin{figure}
\centering
\epsfig{file=edgrid.eps,width=3.5in}
\caption{False Positive Rate, Good Count, Bad Count for $N=10, K \in \{8,9,10\}, 5 \le L \le 15$}
\label{ed-grid}
\end{figure}

\section{Working Application}

While the Minhash signature code provides the raw similarity scores, it is one part of a larger framework composed of several pieces, described below. The input to the process is an XML dump format. (See \url{http://en.wikipedia.org/wiki/Wikipedia:Database_download} for details.) The output of the process is a list of sentence clusters containing the original article title and sentence text. Decomposing the sentence clustering into several pieces allows us to minimize the amount of raw text that gets passed between mappers and reducers, thus improving scalability.

The process can be summarized as follows:

\begin{enumerate}
\item \textbf{Preprocess Wikipedia Pages} - Using a utility from the Cloud9 library, preprocess the raw Wikipedia dump into a SequenceFile format. Working with SequenceFiles gives the option to use block or record compressed input compatible with the MapReduce framework.
\item \textbf{Minhash Wikipedia Sentences} - Reads processed Wikipedia input, and uses a regular expression to match sentences in the article text. Outputs a file of signature ``buckets'' organized by signature. Each output sentence is represented by a (docid,sentenceid) pair.
\item \textbf{Dedup Sentence Buckets} - Because multiple sentences may occur in each bucket, a second MapReduce job is used to map from buckets to raw sentence pairs, removing duplicates.
\item \textbf{Compute Sentence Clusters} - Performs a local (i.e., non-MapReduce) connected component analysis on the dedup-ed pairs to form sentence clusters.
\item \textbf{Get Sentence Text} - Maps clustered sentences (represented by (docid,sentenceid) pairs) back to sentence text by streaming over the Wikipedia input a second time and recovering the corresponding sentences by ID.
\item \textbf{Edit Distance Filter} - \emph{(optional)} For each output cluster and filter all pairs within the cluster by desired edit distance threshold.
\end{enumerate}

\subsection{Limitations to Our Approach}

@@

We performed several analyses to provide more insight into the occurence of duplicate sentences. In particular, we were interested in detecting similar sentences that arose from copying text from one article to another since such instances would represent content that is either redundant, or content that has diverged in the sense that either the original or the copy was edited for clarification of factual correctness and the other was not. Such instances give insight into the Wikipedia editing process, and also represent an opportunity to improve article quality on Wikipedia by identifying places where further editing or article merging may be needed.

@@

>>>>

\title{Similar Sentence Detection Using Locality Sensitive Hashing Technique on Wikipedia}

@@

\author{
\alignauthor Samet Ayhan\\
       \affaddr{University of Maryland}\\
       \affaddr{Dept. of Computer Science}
       \email{sayhan@cs.umd.edu}
\alignauthor Joshua Bradley\\
       \affaddr{University of Maryland}\\
       \affaddr{Dept. of Computer Science}
       \email{jgbrad1@cs.umd.edu}
\alignauthor Sarah E. Weissman\\
       \affaddr{University of Maryland}\\
       \affaddr{College of Information Studies}\\
       \email{sew@umd.edu}
}

\date{October 9, 2013}

\maketitle
\begin{abstract}
The growth of the Internet has enabled collaboration and cooperation on a large scale, resulting in an abundant number of near-duplicate web documents. Information is propagated on a large scale, which means that finding (and correcting) factual errors is made more difficult, since such errors may be duplicated many times over in different locations. Wikipedia articles, which are written and edited by an open community of mostly anonymous users are a prime candidate for copying of information. With few exceptions, the reader of any article can edit the text by copying from another article or an outside source without prior approval or confirmation of correctness. Although Wikipedia has its own policies and procedures for maintaining articles and improving quality, questions of the quality and accuracy of information on the website remain. Although we do not directy attempt to measure the quality of Wikipedia articles, we are interested in finding near-duplicate occurrences at various granularity levels. Such near duplicate sentences may reflect factual information that has been updated in one article and not another, or merely inconsistencies that should be flagged and addressed.

In this paper, we describe a similarity detection algorithm for Wikipedia articles that utilizes a locality sensitive hashing (LSH) technique on the MapReduce framework. The algorithm has been designed and implemented to detect similar content at the sentence level using large Wikipedia dumps. We also describe a heuristic method for classifying types of duplications on Wikipedia, which could be used for selecting and prioritizing editing efforts.
\end{abstract}

@@

\section{Introduction}
The online encyclopedia Wikipedia is a collaboratively edited, multilingual, free resource. As of xxx date, it contains 26 million articles, over 4.2 million in English alone. These articles are written and edited by volunteer contributors around the world. Most articles can be edited by anyone with internet access, and there are currently approximately 100,000 active contributors.

Because of its open nature, Wikipedia has a history of generating controversy over editorial quality and factual correctness. Although some studies have found Wikipedia's accuracy to rival that of traditional encyclopedias, other studies have found numerous factual errors.[References??] Since Wikipedia may be edited anonymously, information may freely copied between web sources and even between Wikipedia articles without verification. Additionally, many articles in Wikipedia lack citations. Although there are active communities of editors who contribute to the upkeep of various articles, often focusing on a partiular subject area and expanding and improving the related content, still much of Wikipedia content is created in an unsupervised manner. Wilkinson found the distribution of article edits on Wikipedia to have a long tail, meaning that a small number of articles have many edits while most articles have few, and that the number of edits is directly related to article quality. Articles with low edits and low editorial attention, are less likely to be updated when new information is available \cite{wilkinson:wiki}.

Under such conditions, this type of large-scale collaborative effort may result in duplicate or near duplicate articles at various granular levels, such as sentence-level or multi-gram granularity \cite{wiki:weblink} [??]. For our purposes, we focus primarily on the sentence level. Near duplicate sentences may reflect factual information that has been updated in one article and not another (for example, information that varies over time, like populations or a country's GDP), or merely inconsistencies that should be flagged and addressed.

Our approach applies minhashin, a well known, scalable, locality sensitive hashing (LSH) technique for near duplicate document detection. Since we are assuming the content we are looking for is not just similar, but has diverged due to an editorial process, as part of this work we investigate how well minhash approximates edit distance.  
We investigate the following questions:
\begin{itemize}[noitemsep,nolistsep]
\item How effectively does minhashing approximate edit distance? Edit distance seems like a reasonable metric for near duplicate sentences, however LSH does not map naturally to the edit distance metric.
\item How well do techniques for detecting near duplicate documents work when applied at the sentence level?
\item Do near duplicate sentences correspond to factual errors?
\end{itemize}

Along with these questions, we also considered the scalability and efficiency issues due to the large size of Wikipedia. To address these issues, we implemented our algorithm on the MapReduce computing framework.

@@

\section{Related Work}
A number of techniques have been implemented to identify similarity between documents in the literature. These techniques usually differ in terms of corpus of interest, signature representing the document, feature-set determined by the system, and the end goal of the analysis.

@@

In order to handle large Wikipedia dumps with the order of tens or hundreds of gigabytes and even a few terabytes, we used MapReduce framework in this work, that was introduced by Dean and Ghemawat \cite{dean:mapreduce}.

@@

In order to measure document similarity we use a common LSH technique known as MinHash (\cite{broder:resemblance}). A minhash signature on a text document is calculated using a parametrized family of hash functions $F_i$, $1 \le i \le N$. (In our case, input ''documents'' are individual sentences found within the content of Wikipedia articles.) Each document is broken up into n-gram ``shingles'' and for each shingle set $S$ a set $\{min_{s \in S}(F_i(s)\}$ of minimum hashes over the hash family is produced. The signature on a document $d$ is represented as a vector of $K$ minhashes, chosen from the set $\{min_{s \in S}(F_i(s)\}$ of minimum hashes. In order to minimize false negatives we apply a technique sometimes known as ``banding'' \cite{ullman:massive} where multiple signatures are produced for each input document.

@@

Our implementation is built on top of the Apache Hadoop MapReduce framework, using utilities from the Cloud9 (\url{https://github.com/lintool/Cloud9}) and WikiClean (\url{https://github.com/lintool/wikiclean/}) libraries for parsing Wikipedia page text from the Wikipedia XML dump format. An open source implementationis available on Github (\url{https://github.com/seweissman/wikiduper}).

%Sentences are parsed out over each document using a regular expression. Sentences that are too large or too small are discarded. The family of hash functions is implemented using a ``Multiply Shift'' (\url{http://en.wikipedia.org/wiki/Universal_hashing}) hashing scheme and is generated from a random seed using the Java Random class. The hash family is also parametrized by hash output key size and this value can be configured to affect the size of the output signatures.

@@

%Jaccard similarity is a measure of set similarity while our stated goal is to find sentences in Wikipedia that have been copied from one article to another and possibly edited. A more natural measure to use for this application is edit distance, but there is no known good locality sensitive hashing family for edit distance \cite{Kirsch06distance-sensitivebloom}.

@@

In order to estimate the affect of shingle length on false positive rate we ran a number of experiments for a 1 G collection of wikipedia documents. For fixed $N=10$ we ran experiments for three values of $K=\{8,9,10\}$ and $8 \le L \le 16$. Setting a normalized edit distance threshold of $e = .25$, we calculated the total number of ''good'' ($e <= .25$) and ''bad'' ($e > .25$) (unique) sentence pairs as well as the false positive rate (\# bad pairs divided by the total number of sentences). Results are presented in Figure \ref{ed-grid}.

As we might expect, as shingle length increases, the false positive rate decreases. However, the total number of good pairs also decreases, implying that our false negative rate is also increasing. When applying these techniques, increased recall will come at the price of decreased precision. Because of this, in a real implementation, it would likely be desirable to set requirements for recall, then use secondary filtering to increase precision. For example, actual edit distance could be applied as a secondary filter.

\begin{figure}
\centering
\epsfig{file=edgridl.eps,width=3.5in}
\caption{False Positive Rate, Good Count, Bad Count for $N=10, K \in \{8,9,10\}, 5 \le L \le 15$}
\label{ed-grid}
\end{figure}

@@

%We performed several analyses to provide more insight into the occurence of duplicate sentences. In particular, we were interested in detecting similar sentences that arose from copying text from one article to another since such instances would represent content that is either redundant, or content that has diverged in the sense that either the original or the copy was edited for clarification of factual correctness and the other was not. Such instances give insight into the Wikipedia editing process, and also represent an opportunity to improve article quality on Wikipedia by identifying places where further editing or article merging may be needed.

@@

\section{Limitations of Our Approach}
